---
title: "Final Project"
author: "Tanay Sawant 19203264"
date: "29/07/2020"
output: html_document
runtime : shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading the required libraries:

```{r, libraries ,warning=FALSE, message=FALSE}

library(dplyr)
library(magrittr)
library(ggplot2)
library(reshape2)
library(ggiraph)
library(gganimate)
library(readr)
library(shiny)
library(tibble)

```

### Question 1.)


```{r, Question 1}

# Reading in the exo planet dataset as tibble.

exo<- as_tibble(read.csv("exo_data.csv"))

# Converting columns of flag and meth as factor. Rest all columns are already as desired in the question.

exo<- exo %>% 
  mutate_at(c("flag","meth"),as.factor)
glimpse(exo[,c(2,14)])

```

### Question 2.)

```{r, Question 2}

# Excluding the exoplanets with an unknown method of discovery.

exo<-filter(exo,meth!="")

```


### Question 3.)


```{r, Question 3, warning=FALSE}

# A Graphic which illustrates the relationship between the 
# log-distances from the Sun and the methods of discovery

p<-ggplot(exo,aes(x=meth,y=log(dist),fill=meth,na.rm=T))+
  geom_boxplot()+ 
  xlab('Method of Discovery') +
  ylab('Log Distance from the Sun') +
  ggtitle('Relationship between Log Distances from the Sun and Method of Discovery')+
  theme(legend.position = "none")
print(p)

```

### Question 4.)


```{r, Question 4, warning=FALSE}

# Adding the Onclick column for the scatterplot below.

exo$onclick=sprintf("window.open(\"%s%s\")",
                         "http://www.openexoplanetcatalogue.com/planet/",
                         as.character(exo$id))

# Scatterplots of the log-mass versus log-distances, separating by methods of discovery

p<-ggplot(exo,
          aes(x=log(mass),y=log(dist),colour=meth))+
  xlab('Planetary Mass (in Log)') +
  ylab('Distance from Sun (in log)') +
  ggtitle('Log Distances from the Sun and Log Planetary Mass values')+
  theme(legend.position = "none")

p<- p + geom_point_interactive(aes(data_id=id,
                                 tooltip = id,
                                 onclick=onclick),
                               size=2)+
  facet_grid(~meth)

ggiraph(code = print(p), width = 0.9)

```



### Question 5.)


```{r, Question 5}

# Renaming the radius into jupiter_radius.

exo<- exo %>%
  rename(jupiter_radius=radius)

# Creating a new column called earth_radius which is 11.2 times the Jupiter radius.

exo<- exo%>%
  mutate(earth_radius=jupiter_radius%>% multiply_by(11.2))

```


### Question 6.)


```{r, Question 6}

# Setting a seed to obtain the same result of kmeans over multiple runs.

set.seed(19203264)

#Focusing on log earth_radius and log period columns have no missing values.

exo<-exo %>%
  filter(!is.na(log(earth_radius)) & !is.na(log(period)))

# Performing kmeans on the two columns with 4 clusters.

kmean=kmeans(log(exo[,c(27,5)]),4)

# Printing the results of the cluster segregation obtained after performing kmeans.

table(kmean$cluster) 

```

### Question 7.)


```{r, Question 7}

# Adding the clustering labels to the dataset through a new factor column called type.
# These labels are according to the image provided on wikipedia.

exo<- exo %>%
  mutate(type = factor(kmean$cluster, labels = c("cold_gas_giants", "others",
                                                 "hot_jupiters","rocky")))

# Checking the corresponding change in the labels of the column type.

table(exo$type)

# Producing the scatterplot highlighting these clusters.

p<-ggplot(exo,
          aes(x=log(period),y=log(earth_radius),colour=type))+
  geom_point()+
  xlab('Period (in Log)') +
  ylab('Earth Radius (in log)') +
  ggtitle('Log Distances from the Sun and Log Planetary Mass values')+
  theme_dark()

print(p)

```

### Question 8.)


```{r, Question 8, warning=FALSE}

# Using a violin plot to illustrate how these clusters relate to the log-mass of the exoplanets.

p<-ggplot(exo,
          aes(x=type, y=log(mass),fill=type))+
  ylab('Planetary Mass (in Log)') +
  xlab('Cluster Type') +
  ggtitle('Log Mass v/s Cluster Type')+
  theme(legend.position = "none")

p<- p + geom_violin()
print(p)
```

### Question 9.)


```{r, Question 9, warning=FALSE}

# Splitting the two columns r_asc and decl into degrees, minutes and seconds for converting them into seconds.

r_asc_split<- colsplit(exo$r_asc, " ", c("degree", "minutes","seconds"))
decl_split<- colsplit(exo$decl, " ", c("hours", "minutes","seconds"))

# Transforming r_asc and decl into two new variables that are the same variables but
# in values of seconds.

exo<- exo%>%
  mutate(r_asc_seconds = r_asc_split$degree %>% multiply_by(3600) %>%
           add(r_asc_split$minutes %>% multiply_by(60)) %>%
           add(r_asc_split$seconds),
         
         decl_seconds = decl_split$hours %>% multiply_by(3600) %>%
           add(decl_split$minutes %>% multiply_by(60)) %>%
           add(decl_split$seconds))

# Plot of Celestial map for the exoplanets.

p<-ggplot(exo,
          aes(x=r_asc_seconds, y=decl_seconds, colour=type))+
  geom_point()+
  xlab('Right Ascension (in Seconds)') +
  ylab('Declination (in Seconds)') +
  ggtitle('Celestial map for the exoplanets')

print(p)

```

### Question 10.)


```{r, Question 10}

# Specific reduced dataset for animated time series.

exo_ts<- exo %>%
  group_by(meth, year) %>%
  summarise(count = n()) %>%
  mutate(count = cumsum(count))

# Animated time series plot.

p<- ggplot(na.omit(exo_ts),
           aes(x=year, y=count, group=meth, col= meth))+
  geom_line(size=0.75)+
  geom_point(size=2.5, col= "black")+
  transition_reveal(year)+
  xlab('Year') +
  ylab('Number of Planets Discovered') +
  ggtitle('Evolution of Number of Planets discovered over the years')+
  theme_dark()+
  theme(axis.text.x=element_text(color = "black", size=11, angle=30, vjust=.8, hjust=0.8))
animate(p)

```

### Question 11 & 14.)


```{r, Question 11 and 14, warning=FALSE}

# Creating shiny.
# The user interface page.

ui <- fluidPage(

#The title to be displayed.
  
  titlePanel("Exoplanet"),

# Sidebar layout format.
  
  sidebarLayout(
    sidebarPanel(

# Slider Input for the Year of Discovery.
      
      sliderInput(inputId = "year",
              label = "Year",
              min = 2009,
              max=2018,
              value = 2009),
  
# select Input for the type of the exoplanet.

      selectInput(inputId = "type",
              label = "Exoplanet Type",
              choices = list("rocky",
                             "hot_jupiters",
                             "cold_gas_giants",
                             "others",
                             "all"),
              selected = "rocky")
    ),


# Main panel.

    mainPanel(
      plotOutput(outputId = "main_plot", height = "300px")
    )
  )
)

# The server page.

server <- function(input, output) {

# Producing the output plot with the given conditions.
    
  output$main_plot <- renderPlot({
    
# If the user selects type as "all" to be displayed on the graph, 
# filtering the data only corresponding to the Year of discovery.
    
    if(input$type=="all")
      x<-exo%>%
        filter(year<=input$year)

# When the user doesn't select type as "all" to be displayed on the graph,
# filtering the data with respect to Year and Type.

    else
      x<-exo%>% 
        filter(year<=input$year,type==input$type)
    
# Scatterplot of log-mass vs log-distance coloured by method of discovery 
# only of the planets that have been discovered till that year.
    
    ggplot(x,
           aes(x=log(mass),y=log(dist),colour=meth))+
      geom_point()+
      xlab('Planetary Mass (in Log)') +
      ylab('Distance from Sun (in log)') +
      ggtitle('Log Distances from the Sun and Log Planetary Mass values')
    
  })
}

# Running the Shiny App.

shinyApp(ui=ui,server=server)

```


### Question 12 & 13.)


```{r, Question 12 and 13}

# Fitting a linear regression model where log period is the response variable and the 
# logs of host_mass, host_temp and axis are the covariates.
# Excluding rows that contain at least one missing value. 
# Including an intercept term in the regression model.

exo_reduced<-exo%>%
  select(period,host_mass,host_temp,axis)%>%
  filter(!is.na(log(period)) & !is.na(log(host_mass)) & !is.na(log(host_temp)) & !is.na(log(axis)))


linear_reg<-lm(log(period)~log(host_mass)+log(host_temp)+log(axis),data = exo_reduced)
summary(linear_reg)

```

We have fitted a linear regression model where **log period** is the response variable and the **log values of host_mass, host_temp and axis** are the covariates. 

The residuals have a range (-1.9,1.2) with the median value being -0.00144.The Residual standard error value is **0.2244** on 616 degrees of freedom.   

The p-value observed in the last line is < 0.05 indicating that our model is a **good fit.** The F-Statistic value observed is 9839 on 3 and 616 DF.  

We observe that the p-values of all the covariates is less than 0.05 indicating that **all the covariates are significant.**    
Specifically,  
A unit change in log value of host_mass causes a decrease in the log period value by 0.26 keeping other covariates constant.  
A unit change in log value of host_temp causes a decrease in the log period value by 0.16 keeping other covariates constant.  
A unit change in log value of axis causes an increase in the log period value by 1.48 keeping other covariates constant.  

In general the standard error of the covariates has to be closer to zero to get a good fit.  
When we observe the standard error values of the covariates, **all the covariates have a standard error value very close to zero which is a good sign.**   

We observe that the **Adjusted R-squared value is 0.9795 i.e approximately 98% of the variation in the log(period) variable is explained by the covariates.**   
It is traditionally known that higher the adjusted R-squared value, better the fit of the model.  
**Overall, we can conclude that our model is a good fit.**